### 2021-02-18 exploratory/exploratory.Rmd
add exploratory rmd

---
title: "Exploratory: Optimal Resolution"
author: "Courtney R Armour"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup_environment, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(ggpubr)
library(nationalparkcolors)
library(data.table)
library(cowplot)

models <- c("rf","glmnet","xgbTree","svmRadial","rpart2")
levels <- c("phylum","class","order","family","genus","otu","asv")
levels_names <- c("Phylum","Class","Order","Family","Genus","OTU","ASV")
pal <- park_palette("Everglades",length(models))
names(pal) <- models
pal2 <- park_palette("Everglades",length(models))
names(pal2) <- c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")

full_df <- NULL
auc_df <- NULL
cv_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))
t_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))

for( m in models ){
  phylum <- read_csv(paste0("./data/final/process/combined-phylum-",m,".csv")) %>% 
    mutate(level="phylum")
  class <- read_csv(paste0("./data/final/process/combined-class-",m,".csv")) %>% 
    mutate(level="class")
  order <- read_csv(paste0("./data/final/process/combined-order-",m,".csv")) %>% 
    mutate(level="order")
  family <- read_csv(paste0("./data/final/process/combined-family-",m,".csv")) %>% 
    mutate(level="family")
  genus <- read_csv(paste0("./data/final/process/combined-genus-",m,".csv")) %>%
    mutate(level="genus")
  otu <- read_csv(paste0("./data/final/process/combined-otu-",m,".csv")) %>%
    mutate(level="otu")
  asv <- read_csv(paste0("./data/final/process/combined-asv-",m,".csv")) %>%
    mutate(level="asv")

  model_df <- bind_rows(phylum,class,order,family,genus,otu,asv)

  full_df <- bind_rows(full_df,model_df)

  m_auc.df <- select(model_df,cv_metric_AUC,AUC,method,level) %>% 
    rename(test_AUC=AUC) %>% 
    pivot_longer(cols = c("cv_metric_AUC","test_AUC"),names_to="auc_type",values_to="AUC")

  m_auc.df$level <- factor(m_auc.df$level,levels=c("phylum","class","order","family","genus","otu","asv"))

  auc_df <- bind_rows(auc_df,m_auc.df)

  cv_med <- m_auc.df %>%
    filter(auc_type == "cv_metric_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC))

  t_med <- m_auc.df %>% 
    filter(auc_type == "test_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC))

  cv_table <- inner_join(cv_table,cv_med,by="level")
  t_table <- inner_join(t_table,t_med,by="level")
}

#compare two models levels within a taxonomic level
p.table.level <- read_csv("./data/analysis/exploratory3/pvalues_by_level.csv")
#compare two taxonomic levels within a model type
p.table.model <- read_csv("./data/analysis/exploratory3/pvalues_by_model.csv")

```