### 2021-08-30 paper/manuscript.tex
update word document

\documentclass[]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{\vspace{-2.5em}}
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\section{Taxonomic Resolution Matters for Microbiome-Based
Classification of Colorectal
Cancer}\label{taxonomic-resolution-matters-for-microbiome-based-classification-of-colorectal-cancer}

\vspace{10mm}

Courtney R. Armour${^1}$, Begüm D.Topçuoğlu, Andrea Garretto${^1}$,
Patrick D. Schloss ${^1}$${^*}$

\vspace{5mm}

${^1}$ Department of Microbiology and Immunology, University of
Michigan, Ann Arbor

\vspace{5mm}

${^*}$ Correspondence: Patrick D. Schloss
\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

\vspace{20mm}

\textbf{observation format - max 1200 words, 2 figures, 25 ref}

\newpage

\subsection{Abstract (max 250 words)}\label{abstract-max-250-words}

\subsection{Importance (max 150 words)}\label{importance-max-150-words}

\newpage

Colorectal cancer is one of the most common cancers in men and women and
a leading cause of cancer related deaths in the United States (1). Early
detection and treatment are essential to increase survival rates, but
for a variety of reasons including the invasiveness and high cost of
screening (i.e.~colonoscopy), many people do not comply with recommended
screening guidelines (2) prompting a need for low cost, non-invasive
detection methods. A growing body of research points to the gut
microbiome as a promising target for non-invasive screening to detect
screen relevant neoplasia (SRNs) consisting of pre-cancerous polyps
(i.e.~advanced adenomas) and carcinomas. Efforts to realize the
diagnostic potential of the gut microbiome in detecting SRNs have
focused on machine learning (ML) methods using abundances of operational
taxonomic unit (OTU) classifications based on amplicon sequencing of the
16S rRNA gene \{\}. However, whether this is the optimal taxonomic
resolution for classifying SRNs from microbiome data is unknown.
Additionally, recent work has pushed for the use of amplicon sequence
variants (ASVs) to replace OTUs for marker-gene analysis because of the
improved resolution with ASVs (3). However, whether the additional
resolution provided by ASVs is useful for ML classification is unclear
\{\}. Since ML classification relies on consistent differences between
groups, its possible that the resolution at the ASV level is too
individualized to accurately differentiate groups. Topçuoğlu \emph{et
al} (4) recently demonstrated how to effectively apply machine learning
(ML) methods to microbiome based classification problems and developed a
framework for applying ML practices in a more reproducible way
(mikropml). This analysis utilizes the reproducible framework developed
by Topçuoğlu \emph{et al} to determine which ML method and taxonomic
level produce the best performing classifier for detecting SRNs from
microbiome data.

Utilizing publicly available 16S rRNA sequence data from stool of
patients with SRNs and healthy controls, we generated abundance tables
with mothur \{\} annotated to phylum, class, order, family, genus, OTU
and ASV levels. Using the taxonomic abundance data, we quantified how
reliably samples could be classified as ``normal'' or ``SRN''
(i.e.~advanced adenoma or carcinoma) using five machine learning methods
with the mikropml R package (methods). Across the five machine learning
methods tested, model performance tended to increase with taxonomic
level usually peaking around genus/OTU level before dropping off
slightly with ASVs (Supplemental Figure 1). Regardless of the taxonomic
level, random forest (RF) models consistently had the largest area under
the receiver operating characteristic curve (AUROC). Within the RF
model, the highest AUROCs were observed for family (mean AUROC: 0.689),
genus (mean AUROC: 0.690), and OTU (mean AUROC: 0.693) level data with
no significant difference between the three (Figure 1A, Supplemental
Figure 2). Performance with ASVs (mean AUROC: 0.676) was significantly
lower than OTUs (p \textless{} 0.01), but comparable to family (p =
0.06) and genus (p = 0.05) levels (Figure 1A). These results suggest
that increased resolution improves model performance until the OTU level
where further taxonomic resolution is not necessarily better for
identifying individuals with SRNs based on microbiome composition.

While comparing AUROC values between models is a useful way to assess
the overall model performance, they summarize the performance across all
thresholds and can be misleading since models with the same AUROC can
have different ROC curve shapes (5). Depending on the intended
implementation of the model, one may want to optimize the true-positive
rate (or sensitivity) over the false-positive rate (or 1-specificity),
or vice versa. In this case, the optimal model will detect as many true
positives (people with SRNs) as possible. To further compare the model
performance across taxonomic levels we compared the sensitivity of the
models at a specificity of 90\%. The highest sensitivity values were
observed for family, genus, and OTU level data (Figure 1B), consistent
with the AUROC results. Phylum, class, order, and ASV sensitivity values
were all significantly lower than family, genus, and OTU sensitivity
values (Figure 1B). This analysis further supports the observation that
finer resolution is does not improve SRN detection.

One hypothesis for the observation that model performance increases from
phylum to OTU level then drops slightly at ASV level is that at higher
taxonomic levels (e.g.~phylum) there are too few taxa and too much
overlap to reliably differentiate between cases and controls. At the
level of genus or OTU there is enough data and variation but at the ASV
level, the data is too specific to individuals and does not overlap
enough. Examination of the prevalence of taxa in samples at each level
supports this idea. A majority of taxa were present in greater than 70\%
of samples at the phylum (67\% of taxa) and class (63\% of taxa) levels.
The opposite was observed at the OTU and ASV level where 50\% and 41\%
of taxa respectively were only present in 20\% or less of the samples
(Supplemental Figure 3). Of note, the ML pipeline includes a
pre-processing step that occurred prior to training and classifying the
ML models. As part of this step, taxa were removed that would not
provide useful information to build the model. For example, perfectly
correlated taxa provide the same information to build the model and thus
can be collapsed. Additionally, features with zero or near-zero variance
across samples would not help the model differentiate groups and thus
can be removed. Interestingly, despite starting with 104,106 ASVs, only
478 (0.5\%) remained after pre-processing. At the OTU level, 705 of the
20,079 OTUs (3.5\%) remained after preprocessing (Table 1). While the
resolution provided by ASVs is useful in certain contexts \{\}, these
results suggest that the resolution is too fine for use in machine
learning classification of SRNs based on microbiome composition.

A look into the most important taxa at each level for classifying
samples revealed some nesting where several genera and their higher
taxonomic classifications were in the top 10 most important taxa
(Supplemental Figure 4). For example, the genus \emph{Gemella} was an
important taxon at the genus and OTU levels and its higher
classifications were also important (\emph{Firmicutes} \textgreater{}
\emph{Bacilli} \textgreater{} \emph{Bacillales} \textgreater{}
\emph{Bacillales Incertae Sedis XI} \textgreater{} \emph{Gemella}).
\emph{Fusobacterium} displayed a similar pattern, except that the family
level classification (\emph{Fusobacteriaceae}) importance was ranked
16th out of 54 families. In the case of unclassified
\emph{Lachnospiraceae}, there were several OTUs with this label that
were in the top 10, however at the genus level this taxon was ranked
lower in importance (21st out of 115 genera) suggesting there may be
some benefit to separating different taxonomic groupings within
\emph{Lachnospiraceae}.

These results demonstrate that consideration of the appropriate
taxonomic resolution for utilizing the microbiome as a predictive tool
is warranted. In general, we found that finer taxonomic resolution
(e.g.~OTU and ASV) did not add additional sensitivity to predicting SRNs
based on microbiome composition. Additionally, at the ASV level the fine
resolution actually impedes model performance due to the sparsity of
shared taxa and led to decreased model performance. The tendancy for ASV
level annotation to split single bacterial genomes into multiple taxa
\{Pat mSphere 2021\} could also be a contributing factor to the sparsity
of shared taxa. Overall, either family, genus, or OTU level taxonomy
appear to perform equally for predicting subjects with SRNs based on the
composition of the gut microbiome. A potential benefit of utilizing
genus or family level data could be that it may allow for merging data
generated from different 16S rRNA gene regions or sequencing platforms.

\subsection{Materials and Methods}\label{materials-and-methods}

\textbf{\emph{Dataset.}} Raw 16S rRNA gene amplicon sequence data
isolated from human gut samples \{Baxter\} was downloaded from NCBI SRA
(accession \#). This dataset contains stool samples from 490 subjects.
Based on the available metadata, samples categorized as normal, high
risk normal, or adenoma were labeled ``normal'' for this analysis and
samples categorized as advanced adenoma or carcinoma were labeled as
``screen relevant neoplasia'' (SRN). This resulted in a total of 261
``normal'' samples and 229 ``SRN'' samples.

\textbf{\emph{Data processing.}} Sequence data was processed with mothur
(1.44.3) using the SILVA reference database (v132) to produce count
tables for phylum, class, order, family, genus, OTU, and ASV following
the Schloss Lab MiSeq SOP described on the mothur website. ASV level
data was also produced using DADA2 to ensure consistent results with a
different pipeline. Data was processed following the DADA2 pipeline, but
setting pool=TRUE to infer ASVs from the whole dataset rather than per
sample. The resulting ASV table was subsampled for consistency with the
mothur data. These results were consistent for ASVs generated with DADA2
(6) (mean AUROC: 0.659).

\textbf{\emph{Machine Learning.}} Machine learning models were run with
the R package mikropml (v0.0.2) \{\} to predict the diagnosis category
(normal vs SRN) of each sample. Data was preprocessed to normalize
values (scale/center), remove values with zero or near-zero variance,
and collapse collinear features using default parameters. Initially the
models were run with default hyperparameters, but were expanded if the
peak performance was at the edge of the hyperparameter range. Each
taxonomic model taxonomic level combination (e.g.~random forest on
genus) was run with 100 different seeds. Each seed split the data into a
training (80\%) and testing (20\%) set, and output performance of the
training and testing as area under the receiver operating curve (AUROC).

To compare performance between taxonomic levels and models, P values
were calculated as previously described \{begum\}. To compare
sensitivity at 90\% specificity, probabilities on the test dataset were
collected for each seed and used to construct ROC curves (R pROC::roc).
From the ROC curves The sensitivity at a specificity of 90\% was pulled
for each seed. An optional output from the mikropml package is the
permuted feature importance which is quantified by iteratively permuting
each feature in the model and assessing the change in model performance.
Features are presumed to be important if the performance of the model,
measured by the AUROC, decreases when that feature is permuted. Ranking
of feature importance was determined by ordering the features based on
the change in AUROC where features with a larger decrease in AUROC are
ranked higher in importance.

To quantify prevalence of the features, the number of samples with
non-zero abundance was divided by the total number of samples resulting
in values ranging from 0 to 1 where 0 indicates the feature is not found
in any samples, 0.5 indicates the feature is found in half of the
samples, an 1 indicating the feature is found in all of the samples.

All code is available at: \textbf{TODO: link to code}

\subsection{Acknowledgements}\label{acknowledgements}

\newpage

\subsection{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figure_1.png}
\end{figure}

\textbf{Figure 1: Random Forest Model Performance.} \textbf{A)} Strip
plot of the area under the receiver operating characteristic curve
(AUROC) values on the test dataset for 100 seeds predicting SRNs using a
random forest model. Black points denote the mean and lines denote the
standard deviation. Dashed line denotes AUROC of 0.5 which is equivalent
to random classification. Significance between taxonomic levels was
quantified by comparing the difference in mean AUROC and is denoted by
letters A through E on the right side of the plot; taxonomic levels with
the same letter are in the same significance group and are not
significantly different from one another. \textbf{B)} Strip plot of the
sensitivity at a specificity of 90\% across the 100 model iterations for
each taxonomic level. Black points denote the mean and the lines denote
the standard deviation. The letters W through Z on the right side of the
plot denote the significance groups.

\newpage

\subsection{Tables}\label{tables}

\begin{longtable}[c]{@{}lrrc@{}}
\toprule\addlinespace
Taxonomic Level & Number of Features &
\makecell[c]{Number of Features \\ After Preprocessing} &
\makecell[c]{Percent of Features Kept \\ After Preprocessing}
\\\addlinespace
\midrule\endhead
Phylum & 19 & 9 & 47.4 \%
\\\addlinespace
Class & 36 & 19 & 52.8 \%
\\\addlinespace
Order & 65 & 28 & 43.1 \%
\\\addlinespace
Family & 124 & 54 & 43.5 \%
\\\addlinespace
Genus & 316 & 115 & 36.4 \%
\\\addlinespace
OTU & 20,079 & 705 & 3.5 \%
\\\addlinespace
ASV & 104,106 & 478 & 0.5 \%
\\\addlinespace
\bottomrule
\end{longtable}

\textbf{Table 1: Summary of Features.} Overview of the number of
features at each taxonomic level before and after preprocessing as
described in the methods.

\newpage

\subsection{Supplemental Figures}\label{supplemental-figures}

\begin{figure}[htbp]
\centering
\includegraphics{figure_s1.png}
\end{figure}

\textbf{Supplemental Figure 1: Model Performance across Taxonomy.}
Boxplots of AUROC values on the test dataset split for 100 seeds for
each model type across all taxonomic levels.

\newpage

\includegraphics{figure_s2.png}\{height=``50\%''\}

\textbf{Supplemental Figure 2: Averaged ROC curves} ROC curves with
averaged true positive rate across the 100 iterations of the model. The
shaded region represents the standard deviation form the mean.Dashed
line represents an AUROC of 0.5, which is equivalent to random
classification. The mean AUROC for each taxonomic level is printed on
the bottom right of the plot.

\newpage

\includegraphics{figure_s3.png}\{height=``45\%''\}\\\textbf{Supplemental
Figure 3: Prevalence of Taxa in Samples.} Distribution of the prevalence
of taxa across samples at each taxonomic level. Percent of samples is
split into 10 groups where the first is taxa present in 0 to 10\% of
samples, then \textgreater{}10\% to 20\% of samples, and so on. The
total number of taxa for each taxonomic level after preprocessing is in
parenthesis next to the title of the plot (the name of the taxonomic
level).

\newpage

\includegraphics{figure_s4.png}\{height=``85\%''\}\\\textbf{Supplemental
Figure 4: Top 10 important taxa at each taxonomic level.} Summary of the
10 most important taxa for the random forest models at each taxonomic
level based on the average decrease in AUROC when the feature is
permuted.

\newpage

\subsection{References}\label{references}

1. \textbf{{Siegel} R}, \textbf{{Miller} K}, \textbf{{Sauer} A},
\textbf{{Fedewa} S}, \textbf{{Butterly} L}, \textbf{{Anderson} J},
\textbf{{Cercek} A}, \textbf{{Smith} R}, \textbf{{Jemal} A}. 2020.
Colorectal cancer statistics, 2020. CA: A Cancer Journal for Clinicians
\textbf{70}:145--164.
doi:\href{http://dx.doi.org/https://doi.org/10.3322/caac.21601}{https://doi.org/10.3322/caac.21601}.

2. \textbf{{García} G}, \textbf{{Z} A}. 2011. Factors influencing
colorectal cancer screening participation. Gastroenterology Research and
Practice \textbf{2012}:e483417.
doi:\href{http://dx.doi.org/10.1155/2012/483417}{10.1155/2012/483417}.

3. \textbf{{Callahan} B}, \textbf{{McMurdie} P}, \textbf{{Holmes} S}.
2017. Exact sequence variants should replace operational taxonomic units
in marker-gene data analysis. The ISME Journal \textbf{11}:2639--2643.
doi:\href{http://dx.doi.org/10.1038/ismej.2017.119}{10.1038/ismej.2017.119}.

4. \textbf{{Topçuo{ğ}lu} B}, \textbf{{Lesniak} N}, \textbf{{Ruffin} M},
\textbf{{Wiens} J}, \textbf{{Schloss} P}. 2020. A framework for
effective application of machine learning to microbiome-based
classification problems. mBio \textbf{11}.
doi:\href{http://dx.doi.org/10.1128/mBio.00434-20}{10.1128/mBio.00434-20}.

5. \textbf{{Lobo} J}, \textbf{{Jiménez-Valverde} A}, \textbf{{Real} R}.
2008. AUC: a misleading measure of the performance of predictive
distribution models. Global Ecology and Biogeography
\textbf{17}:145--151.
doi:\href{http://dx.doi.org/10.1111/j.1466-8238.2007.00358.x}{10.1111/j.1466-8238.2007.00358.x}.

6. \textbf{{Callahan} B}, \textbf{{McMurdie} P}, \textbf{{Rosen} M},
\textbf{{Han} A}, \textbf{{Johnson} A}, \textbf{{Holmes} S}. 2016.
DADA2: High-resolution sample inference from illumina amplicon data.
Nature Methods \textbf{13}:581--583.
doi:\href{http://dx.doi.org/10.1038/nmeth.3869}{10.1038/nmeth.3869}.

\end{document}